---
title: "DPP Reader Paper Analyses"
author: "Marie Ozanne"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(coda)
library(data.table)
library(reshape2)
library(dplyr)
library(egg)
library(ggplot2)
library(ggforce)
library(R2OpenBUGS)
library(xtable)
```

# Preliminary work

```{r echo=FALSE}
dpp_all <- read.csv("G:/My Drive/Research/Sensitivity_Specificity/Data/DPPReaderData_cleaned.csv")
head(dpp_all)
summary(dpp_all)

names(dpp_all)[1:9] <- c(names(dpp_all)[1:4], "PCR_avg_CT", "PCR", "DPP_reader", "DPP", "Clinical_status")

dpp_all <- dpp_all %>% select(ID, Sex, Age, PCR, DPP_reader, DPP, Clinical_status)
summary(dpp_all)
dpp_all[which(is.na(dpp_all$DPP_reader)),]

## One dog with missing DPP and DPP reader (ID: Penelope)
## Will omit 1 observation
dpp_usable <- na.omit(dpp_all) ## left with 1384 observations

str(dpp_usable)
```

## Identifying other missing values

```{r}
## Investigating other missing values for demographic variables ("")
### Sex
unique(dpp_usable$Sex)
# dpp_usable[which(dpp_usable$Sex==""),]

# dpp_usable[which(dpp_usable$ID %in% dpp_usable[which(dpp_usable$Sex==""),]$ID),]

## Add sex for Gumbo:
dpp_usable[dpp_usable$ID=="Gumbo",]$Sex <- "F"

## Print others for which sex is missing
# dpp_usable[which(dpp_usable$ID %in% dpp_usable[which(dpp_usable$Sex==""),]$ID),]

### Age
unique(dpp_usable$Age)
# dpp_usable[which(dpp_usable$Age==""|dpp_usable$Age=="."),]

## Look at PCR and DPP
unique(dpp_usable$PCR) ## What should I do with BL?
dpp_usable$PCR <- tolower(dpp_usable$PCR)
unique(dpp_usable$DPP)
dpp_usable$DPP <- tolower(dpp_usable$DPP)

## Print all dogs for which sex or age is missing
dpp_usable[which(dpp_usable$Age==""|dpp_usable$Age=="."|dpp_usable$Sex=="" | dpp_usable$PCR=="bl"),]
```

### Proceed with BL PCR observations designated as negative

This is the preference Breanna had (3/16/2021).

```{r}
dpp_BLneg <- dpp_usable
dpp_BLneg[dpp_BLneg$PCR == "bl",]$PCR <- "negative"
summary(dpp_BLneg)

## Some plots
dpp_BLneg$diag_pos <- 1*(dpp_BLneg$PCR=="positive" | dpp_BLneg$DPP=="positive")

tmp <- dpp_BLneg %>%
  group_by(Sex, Age) %>%
  summarize(
    emp_odds_diag_pos = mean(diag_pos==1)/(1-mean(diag_pos==1))
  )
```

### Proceed with BL PCR observations designated as positive

We are going to run the models both ways to see how sensitive results are to this choice.

```{r}
dpp_BLpos <- dpp_usable
dpp_BLpos[dpp_BLpos$PCR == "bl",]$PCR <- "positive"
summary(dpp_BLpos)

## Some plots
dpp_BLpos$diag_pos <- 1*(dpp_BLpos$PCR=="positive" | dpp_BLpos$DPP=="positive")

tmp <- dpp_BLpos %>%
  group_by(Sex, Age) %>%
  summarize(
    emp_odds_diag_pos = mean(diag_pos==1)/(1-mean(diag_pos==1))
  )
```

## Other summary plots

```{r}

ggplot(tmp, aes(x=as.numeric(Age), y=emp_odds_diag_pos, color=Sex)) + 
  geom_point() + 
  theme_bw()

ggplot(dpp_usable, aes(x=as.numeric(Age), y=DPP_reader, color=Sex)) + 
  geom_point() + 
  theme_bw()

ggplot(dpp_usable, aes(x=DPP_reader, color=Sex)) + 
  geom_density(size=1) + 
  theme_bw()

dpp_usable <- dpp_usable %>%
  mutate(
    log_DPP_reader = log(DPP_reader + 0.1)
  )

ggplot(dpp_usable, aes(x=log_DPP_reader, color=Sex)) + 
  geom_density(size=1) + 
  theme_bw()

dpp_usable %>% group_by(Sex) %>%
  summarize(
    mean_log_DPP_reader = mean(log_DPP_reader),
    sd_log_DPP_reader = sd(log_DPP_reader)
  )

ggplot(dpp_usable, aes(x=as.numeric(Age), y=log_DPP_reader, color=Sex)) + 
  geom_point() + 
  theme_bw() + 
  geom_smooth(method="lm")

```



# Model 1: Dichotomized DPP and PCR

## MCMC via OpenBUGS (no transformation)

```{r Open Bugs Model ContDPP setup}
## ranges of sensitivities and specificities
sens.pcr.range <- c(0.839, 0.990)
sens.dpp.range <- c(0.832, 0.930)
spec.pcr.range <- c(0.871, 0.970)
spec.dpp.range <- c(0.682, 0.951)

## ranges of sensitivities and specificities
sens.pcr <- mean(sens.pcr.range) #0.999
sens.dpp <- mean(sens.dpp.range) #0.001
spec.pcr <- mean(spec.pcr.range) #0.999
spec.dpp <- mean(spec.dpp.range) #0.001

## range of prevalence for visceral leishmaniasis
prev.range <- c(0.05, 0.10)
prev <- mean(prev.range)
```

## Implementation: Dichotomized Tests

### BL PCR designated as negative

```{r Open Bugs Model 1a, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(dpp_BLneg$PCR_positive=="positive")
y.dpp <- 1*(dpp_BLneg$DPP_positive=="positive")
  
nind <- nrow(dpp_BLneg)
  
Sex2 <- 1*(dpp_BLneg$Sex=="M")
# Age <- as.numeric(dpp_BLneg$Age)-mean(as.numeric(dpp_BLneg$Age))
Age <- as.numeric(dpp_BLneg$Age)-3 ## most common age in sample
S.a <- sens.pcr/(1-sens.pcr)
Sp.a <- spec.pcr/(1-spec.pcr)
S.a_dpp <- sens.dpp/(1-sens.dpp)
Sp.a_dpp <- spec.dpp/(1-spec.dpp)
  
data <- list(nind=nind,
             y.pcr=y.pcr,
             y.dpp=y.dpp,
             Sex=Sex2,
             Age=Age,
             S.a=S.a,
             Sp.a=Sp.a,
             S.a_dpp=S.a_dpp, ## function of mean sensitivity for DPP
             Sp.a_dpp=Sp.a_dpp ## function of mean specificity for DPP
             )
  
## Initialize all prior values
inits <- function(){
    list(S.pcr=rbeta(1,S.a,1),
         Sp.pcr=rbeta(1,Sp.a,1),
         S.dpp=rbeta(1,S.a_dpp,1),
         Sp.dpp=rbeta(1,Sp.a_dpp,1),
         B0=rnorm(0,1),
         B1=rnorm(0,1),
         B2=rnorm(0,1),
         B3=rnorm(0,1),
         eps=rnorm(nind,0,5000))
}
 
## --- No hierarchical centering --- ## 
## Start MCMC simulation
sens_spec.sim_m1_BLneg <- bugs(data=data, 
                          inits=inits, 
                          model.file="Model1b_separate_tests_GB.txt",
                          parameters=c("B0",
                                       "B1",
                                       "B2",
                                       # "B3",
                                       "S.pcr",
                                       "Sp.pcr",
                                       "S.dpp",
                                       "Sp.dpp",
                                       "pi.D",
                                       "eps"),
                          n.chains=3, n.iter=40000, n.burnin=20000, 
                          codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_m1_BLneg <- read.bugs(sens_spec.sim_m1_BLneg)
model1_BLneg_df <- do.call(rbind.data.frame, codaobject_m1_BLneg)

save(model1_BLneg_df, file="C:/Users/mozanne/Documents/Research/DPP_LCA/Models/Model Fits/model1_BLneg_post_prec1.Rdata")
```