---
title: "Latent Class Models for CVL"
subtitle: "Continuous DPP, Dichotomous Assay PCR"
author: "Marie Ozanne"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(coda)
library(data.table)
library(reshape2)
library(dplyr)
library(egg)
library(ggplot2)
library(ggforce)
library(R2OpenBUGS)
library(xtable)
```

## Data Summary and Transformations

```{r fig.align='center'}
## Read in data
ss_data2 <- read.csv(file="G:/My Drive/Research/Sensitivity_Specificity/Data/clean_ss_data2.csv",
                     header = TRUE)

## Add subject index to data frame (for nested random effects)
ss_data2$subject <- group_indices(ss_data2, ID)

## Plot DPP
head(ss_data2)

h1 <- ggplot(data=ss_data2, aes(x=DPP, fill=PCR_positive)) + 
  geom_histogram(bins=50) +
  geom_vline(xintercept=10, linetype="dashed") + 
  theme_bw() + 
  xlab("DPP Reader")
h1

### Add epsilon to only DPP of 0 (147)
ss_data2$DPP[147] <- ss_data2$DPP[147] + 0.01

## sqrt transformation for DPP
h2 <- ggplot(data=ss_data2, aes(x=sqrt(DPP), fill=PCR_positive)) + 
  geom_histogram(bins=50) +
  geom_vline(xintercept=sqrt(10), linetype="dashed") + 
  theme_bw() + 
  xlab("DPP Reader")
h2

## log transformation for DPP
h3 <- ggplot(data=ss_data2, aes(x=log(DPP), fill=PCR_positive)) + 
  geom_histogram(bins=50) +
  geom_vline(xintercept=log(10), linetype="dashed") + 
  theme_bw() + 
  xlab("DPP Reader")
h3

## Summarize data
ss_data2 %>% 
  group_by(DPP_positive2, PCR_positive) %>%
  summarise(Count=n(),
            pct_Male=mean(Sex=="M"),
            median_age=median(Age),
            quantile.025_age=quantile(Age, c(0.025)),
            quantile.975_age=quantile(Age, c(0.975)),
            median_DPP=median(DPP),
            quantile.025_DPP=quantile(DPP, c(0.025)),
            quantile.975_DPP=quantile(DPP, c(0.975)))
```

## Model



## MCMC via OpenBUGS (no transformation)

```{r Open Bugs Model ContDPP setup}
## ranges of sensitivities and specificities
sens.pcr.range <- c(0.839, 0.990)
sens.dpp.range <- c(0.832, 0.930)
spec.pcr.range <- c(0.871, 0.970)
spec.dpp.range <- c(0.682, 0.951)

## ranges of sensitivities and specificities
sens.pcr <- mean(sens.pcr.range) #0.999
sens.dpp <- mean(sens.dpp.range) #0.001
spec.pcr <- mean(spec.pcr.range) #0.999
spec.dpp <- mean(spec.dpp.range) #0.001

## range of prevalence for visceral leishmaniasis
prev.range <- c(0.05,0.10)
prev <- mean(prev.range)
```

```{r}
## Prior information - mixture of Gammas
m1 <- 150.3 #mean DPP for symptomatic Leish (from dogs with 3+ clinical signs)
v1 <- 103.7^2 #DPP variance for symptomatic Leish (from dogs with 3+ clinical signs)
h_a1 <- m1^2/v1 #shape parameter for prior DPP score for Y|D=1
h_b1 <- m1/v1 #rate parameter for prior DPP score for Y|D=1

m0 <- 2 #mean DPP for negative Leish (less sure about these)
v0 <- 1 #DPP variance for negative Leish 
h_a0 <- m0^2/v0
h_b0 <- m0/v0


```

## MCMC via OpenBUGS (with no transformation)



## Results (no transformation)

```{r}
## Read in saved data
# load("G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/model2_posterior.Rdata")
```

## MCMC via OpenBUGS (with Sqrt transformation)

### Set hyperpriors on mixture of Gammas

```{r gamma hyperprior specification sqrt trans}
## Prior information - mixture of Gammas (with sqrt transformation)
m1 <- mean(sqrt(ss_data2[ss_data2$ClinicalStatus=="S",]$DPP)) #mean DPP for symptomatic Leish (from dogs with 3+ clinical signs)
v1 <- var(sqrt(ss_data2[ss_data2$ClinicalStatus=="S",]$DPP)) #DPP variance for symptomatic Leish (from dogs with 3+ clinical signs)
h_a1 <- m1^2/v1 #shape parameter for prior DPP score for Y|D=1
h_b1 <- m1/v1 #rate parameter for prior DPP score for Y|D=1

m0 <- mean(sqrt(ss_data2[ss_data2$ClinicalStatus=="N",]$DPP)) #mean DPP for negative Leish (less sure about these)
v0 <- var(sqrt(ss_data2[ss_data2$ClinicalStatus=="N",]$DPP)) #DPP variance for negative Leish 
h_a0 <- m0^2/v0
h_b0 <- m0/v0
```

### Run through OpenBUGS

```{r Open Bugs Model ContDPP sqrt trans, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp_c <- sqrt(ss_data2$DPP)
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
Ind_Age <- 1*(Age >= 6)
S.a <- sens.pcr/(1-sens.pcr)
Sp.a <- spec.pcr/(1-spec.pcr)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp_c=y.dpp_c,
               Sex=Sex2,
               Age=Age,
               # Ind_Age=Ind_Age,
               h_a0=1/h_a0, # to get the mean to be h_a0
               h_a1=1/h_a1,
               h_b0=1/h_b0,
               h_b1=1/h_b1,
               S.a=S.a, ## function of mean sensitivity for PCR
               Sp.a=Sp.a ## function of mean specificity for PCR
             )

## Initialize prior values
inits <- function(){
    list(B0=rnorm(0,1),
         B1=rnorm(0,1),
         B2=rnorm(0,1),
         B3=rnorm(0,1),
         a0=rexp(4),
         b0=rexp(1),
         eps=rnorm(nind,0,5000)
         )
}

## MCMC Simulation
sens_spec.sim_contDPP_sqrt <- bugs(data=data, 
                              inits=inits, 
                              model.file="Model_cDPP_mixGamma.txt", ##Variables missing from the model
                              parameters=c("B0",
                                           "B1",
                                           "B2",
                                           "B3",
                                           "S.pcr",
                                           "Sp.pcr",
                                           "a1",
                                           "b1",
                                           "a0",
                                           "b0",
                                           "pi.D",
                                           "D" 
                                           # "eps"
                                           ),
                              n.chains=3, n.iter=30000, n.burnin=10000, 
                              codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_contDPP_sqrt <- read.bugs(sens_spec.sim_contDPP_sqrt)
model1_contDPP_sqrt <- do.call(rbind.data.frame, codaobject_contDPP_sqrt)

save(model1_contDPP_sqrt, file="G:/My Drive/Research/Sensitivity_Specificity/Models/Model Fits/model2_posterior_sqrt.Rdata")

```

### Checking convergence

- Check trace plots
- Check Gelman-Rubin statistic (all are close to 1 for basis parameters)

```{r}
model1_contDPP_sqrt$itr <- rep(1:20000, times=3)
model1_contDPP_sqrt$chain <- rep(1:3, each=20000)

par(mfrow=c(2,2))
## Graphial summaries of posterior distributions
hist(exp(model1_contDPP_sqrt$B0)/(1+exp(model1_contDPP_sqrt$B0)), main="prevalence", xlab="posterior draws")
abline(v=mean(exp(model1_contDPP_sqrt$B0)/(1+exp(model1_contDPP_sqrt$B0))), lty="dashed", col="red")

hist(model1_contDPP_sqrt$B0, main="prevalence", xlab="posterior draws")
abline(v=mean(model1_contDPP_sqrt$B0), lty="dashed", col="red")

hist(model1_contDPP_sqrt$B1, main="Sex", xlab="posterior draws")
abline(v=mean(model1_contDPP_sqrt$B1), lty="dashed", col="red")
# hist(model1_contDPP_sqrt$B2, main="Age", xlab="posterior draws")
# abline(v=mean(model1_contDPP_sqrt$B2), lty="dashed", col="red")
# hist(model1_contDPP_sqrt$B3, main="Age/Sex Interaction", xlab="posterior draws")
# abline(v=mean(model1_contDPP_sqrt$B3), lty="dashed", col="red")

## Trace plots for logistic regression parameters
p0 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=B0, col=as.factor(chain))) +
             geom_line()
p1 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=B1, col=as.factor(chain))) +
             geom_line()
p2 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=B2, col=as.factor(chain))) +
             geom_line()
p3 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=B3, col=as.factor(chain))) +
             geom_line()

ggarrange(p0,p1,p2,p3, ncol=2)

## Calculate Gelman-Rubin diagnostic statistic
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$B0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$B0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$B0)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$B1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$B1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$B1)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$B2),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$B2),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$B2)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$B3),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$B3),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$B3)))

## Trace plots for mixture of gamma parameters and sensitivity/specificity for PCR
p4 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=a1, col=as.factor(chain))) +
             geom_line()
p5 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=b1, col=as.factor(chain))) +
             geom_line()
p6 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=a0, col=as.factor(chain))) +
           geom_line()
p7 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=b0, col=as.factor(chain))) +
             geom_line()
p8 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=S.pcr, col=as.factor(chain))) +
             geom_line()
p9 <- ggplot(data=model1_contDPP_sqrt, aes(x=itr, y=Sp.pcr, col=as.factor(chain))) +
             geom_line()

ggarrange(p4, p5, p6, p7, p8, p9, ncol=2)

## Calculate Gelman-Rubin diagnostic statistic
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$a1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$a1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$a1)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$b1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$b1),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$b1)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$a0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$a0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$a0)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$b0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$b0),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$b0)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$S.pcr),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$S.pcr),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$S.pcr)))
gelman.diag(mcmc.list(mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==1,]$Sp.pcr),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==2,]$Sp.pcr),
                      mcmc(model1_contDPP_sqrt[model1_contDPP_sqrt$chain==3,]$Sp.pcr)))
```

### Results summary - regression coefficients (sqrt transformation)

```{r posterior table sqrt transformation, echo=FALSE}
## load(file="G:/My Drive/Research/Sensitivity_Specificity/Models/Model Fits/model2_posterior_sqrt.Rdata")

tmp2 <- model1_contDPP_sqrt[,names(model1_contDPP_sqrt) %in% c("B0","B1","B2","B3","S.pcr","Sp.pcr","a1","b1","a0","b0")]

Pg0_fcn <- function(x){
  mean(x > 0)
}

tmp2_summary <- data.frame(parameter=names(tmp2),
                           mean=apply(tmp2, 2, mean),
                           median=apply(tmp2, 2, median),
                           LB=apply(tmp2, 2, quantile, 0.025),
                           UB=apply(tmp2, 2, quantile, 0.975),
                           Pg0=apply(tmp2, 2, Pg0_fcn))
xtable(tmp2_summary)
```

### Results summary - disease state classification population (sqrt transformation)

```{r}
invlogit <- function(B0, B1, B2, B3, Sex, Age){ #Sex=1 if male
  # exp(B0+B1*Sex+B2*Age)/(1+exp(B0+B1*Sex+B2*Age))
  exp(B0+B1*Sex+B2*Age+B3*Sex*Age)/(1+exp(B0+B1*Sex+B2*Age+B3*Sex*Age))
}

## logistic curves: Model cont DPP - sqrt transformation
post_df_M2 <- data.frame(Age=rep(1:14,2),
                         Sex=c(rep("M",14),rep("F",14)))
post_df_M2$mean_prob <- invlogit(B0=tmp2_summary$mean[1], 
                                 B1=tmp2_summary$mean[2],
                                 B2=tmp2_summary$mean[3],
                                 B3=tmp2_summary$mean[4],
                                 Sex=1*(post_df_M2$Sex=="M"),
                                 Age=post_df_M2$Age) 
post_df_M2$LB_prob <- invlogit(B0=tmp2_summary$LB[1], 
                                 B1=tmp2_summary$LB[2],
                                 B2=tmp2_summary$LB[3],
                                 B3=tmp2_summary$LB[4],
                                 Sex=1*(post_df_M2$Sex=="M"),
                                 Age=post_df_M2$Age) 
post_df_M2$UB_prob <- invlogit(B0=tmp2_summary$UB[1], 
                                 B1=tmp2_summary$UB[2],
                                 B2=tmp2_summary$UB[3],
                                 B3=tmp2_summary$UB[4],
                                 Sex=1*(post_df_M2$Sex=="M"),
                                 Age=post_df_M2$Age)

l1_prob <- ggplot(post_df_M2, aes(x=Age, y=mean_prob, color=Sex)) + 
  geom_point() + 
  geom_line() + 
  geom_ribbon(aes(ymin=LB_prob, ymax=UB_prob, color=Sex, fill=Sex), alpha=0.2) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(fill = "Sex") +
  ylab("P(D=1|X)") + 
  facet_zoom(xlim=c(1,4))
l1_prob
# ggsave(l1_prob, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_1_invlogit.eps")
# ggsave(l1_prob, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_1_invlogit.pdf")
# ggsave(l1_prob, file="G:/My Drive/JSM/JSM 2020/Model_1_invlogit.pdf")

```

# Results summary -- individual classification (all DPP transformations)

```{r classification figures model results, echo=FALSE}
fig_df <- data.frame(mean_piD_M1=apply(model1_contDPP[,grep("pi.D",
                                                       names(model1_contDPP))],
                                       2, median),
                     mean_piD_M2=apply(model1_contDPP_sqrt[,grep("pi.D",
                                                            names(model1_contDPP_sqrt))],
                                       2, median),
                     sd_piD_M1=apply(model1_contDPP[,grep("pi.D",
                                                     names(model1_contDPP))],
                                     2, sd),
                     sd_piD_M2=apply(model1_contDPP_sqrt[,grep("pi.D",
                                                          names(model1_contDPP_sqrt))],
                                     2, sd),
                     lb_piD_M1=apply(model1_contDPP[,grep("pi.D",
                                                     names(model1_contDPP))],
                                     2, quantile, probs=0.1),
                     lb_piD_M2=apply(model1_contDPP_sqrt[,grep("pi.D",
                                                          names(model1_contDPP_sqrt))],
                                     2, quantile, probs=0.1),
                     ub_piD_M1=apply(model1_contDPP[,grep("pi.D",
                                                     names(model1_contDPP))],
                                     2, quantile, probs=0.9),
                     ub_piD_M2=apply(model1_contDPP_sqrt[,grep("pi.D",
                                                          names(model1_contDPP_sqrt))],
                                     2, quantile, probs=0.9),
                     DPP=ss_data2$DPP,
                     Age=ss_data2$Age,
                     Sex=ss_data2$Sex,
                     PCR=ss_data2$PCR_positive,
                     ClinicalStatus=ss_data2$ClinicalStatus,
                     ID=ss_data2$ID)

p1 <- ggplot(data=fig_df) + 
  geom_point(aes(x=DPP, y=mean_piD_M1, color=as.factor(PCR), pch=ClinicalStatus)) +
  theme_bw() + 
  theme(legend.position = "none") +
  ylab("P(D=1|X)") +
  xlab("DPP Reader Score") +
  labs(color="PCR Status") +
  geom_vline(xintercept = 10, lty="dashed", color="gray") +
  geom_hline(yintercept = 0.5, lty="dashed", color="gray") +
  geom_errorbar(aes(ymin=lb_piD_M1, ymax=ub_piD_M1,
                    x=DPP, y=mean_piD_M1), width=.2) + 
  # ylim(0,1) +
  facet_zoom(xlim=c(0,10))

p2 <- ggplot(data=fig_df, ) + 
  geom_point(aes(x=DPP, y=mean_piD_M2, color=as.factor(PCR), pch=ClinicalStatus)) +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  labs(pch="Clinical Status") +
  # ylab("P(D=1|X)") +
  ylab("") +
  xlab("DPP Reader Score") +
  labs(color="PCR Status") +
  geom_vline(xintercept = 10, lty="dashed", color="gray") +
  geom_hline(yintercept = 0.5, lty="dashed", color="gray") +
  geom_errorbar(aes(ymin=lb_piD_M2, ymax=ub_piD_M2,
                   x=DPP, y=mean_piD_M2), width=.2) +
  # ylim(0,1) +
  facet_zoom(xlim=c(0,10))
p1
# ggsave(p1, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_1_ind_D.eps")
# ggsave(p1, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_1_ind_D.tiff")
# ggsave(p1, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_1_ind_D.pdf")
p2
# ggsave(p2, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_2_ind_D.eps")
# ggsave(p2, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_2_ind_D.tiff")
# ggsave(p2, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/Model_2_ind_D.pdf")
# ggarrange(p1, p2, nrow=1)
ggarrange(p1, p2, nrow=1)
# ggsave(p3, file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/M1_M2_ind_D.eps")
# ggsave(ggarrange(p1, p2, nrow=1), file="G:/My Drive/Research/Sensitivity_Specificity/Paper Draft/M1_M2_ind_D.pdf")

## Look at the individuals that have posterior intervals wider than 0.7
summary(fig_df[fig_df$ub_piD_M2-fig_df$lb_piD_M2 > 0.7,names(fig_df) %in% c("ID", "DPP", "Age", "Sex", "PCR", "ClinicalStatus")])
summary(fig_df[fig_df$ub_piD_M2-fig_df$lb_piD_M2 <= 0.7,names(fig_df) %in% c("ID", "DPP", "Age", "Sex", "PCR", "ClinicalStatus")])
```


## MCMC via OpenBUGS (constraint on ordering of Gamma distributions)

```{r}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp_c <- sqrt(ss_data2$DPP)
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
S.a <- sens.pcr/(1-sens.pcr)
Sp.a <- spec.pcr/(1-spec.pcr)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp_c=y.dpp_c,
               Sex=Sex2,
               Age=Age,
               # b1=0.01,
               # b0=1.82,
               S.a=S.a, ## function of mean sensitivity for PCR
               Sp.a=Sp.a ## function of mean specificity for PCR
             )

## Initialize prior values
inits <- function(){
    list(B0=rnorm(0,1),
         B1=rnorm(0,1),
         B2=rnorm(0,1),
         B3=rnorm(0,1),
         a0=rgamma(0.01, 0.01), #3.70,
         b0=rgamma(0.01, 0.01), #1.82, 
         a1=rgamma(0.01, 0.01), #0.52, 
         b1=rgamma(0.01, 0.01), #0.01,
         eps=rnorm(nind,0,5000)
         )
}

## MCMC Simulation
sens_spec.sim_contDPP_oc <- bugs(data=data, 
                              inits=inits, 
                              model.file="Model_cDPP_mixGamma_orderingconstraint.txt", 
                              parameters=c("B0",
                                           "B1",
                                           "B2",
                                           "B3",
                                           "S.pcr",
                                           "Sp.pcr",
                                           "a1",
                                           "b1",
                                           "a0",
                                           "b0",
                                           "pi.D",
                                           "eps"
                                           ),
                              n.chains=3, n.iter=30000, n.burnin=10000, 
                              codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_contDPP_oc <- read.bugs(sens_spec.sim_contDPP_oc)
model1_contDPP_oc <- do.call(rbind.data.frame, codaobject_contDPP_oc)

save(model1_contDPP_oc, file="G:/My Drive/Research/Sensitivity_Specificity/Models/Model Fits/model2_posterior_oc.Rdata")
```


## MCMC via OpenBUGS (log transformation)

### Set hyperpriors on mixture of Gaussians

```{r Gaussian hyperprior specification log trans}
## Prior information - mixture of Gaussians (with log transformation)
m1 <- mean(log(ss_data2[ss_data2$ClinicalStatus=="S",]$DPP)) #mean DPP for symptomatic Leish (from dogs with 3+ clinical signs)
v1 <- var(log(ss_data2[ss_data2$ClinicalStatus=="S",]$DPP)) #DPP variance for symptomatic Leish (from dogs with 3+ clinical signs)

m0 <- mean(log(ss_data2[ss_data2$ClinicalStatus=="N",]$DPP)) #mean DPP for negative Leish (less sure about these)
v0 <- var(log(ss_data2[ss_data2$ClinicalStatus=="N",]$DPP)) #DPP variance for negative Leish

h_mu1 <- m1
h_mu0 <- m0
h_p1 <- 1/v1 ##OpenBUGS parameterized in terms of precision
h_p0 <- 1/v0
```

### Run through OpenBUGS

```{r Open Bugs Model ContDPP log trans, include=FALSE, message=FALSE}
## Specify data
y.pcr <- 1*(ss_data2$PCR_positive=="Positive")
y.dpp_c <- log(ss_data2$DPP)
  
nind <- nrow(ss_data2)
  
Sex2 <- 1*(ss_data2$Sex=="M")
Age <- as.numeric(ss_data2$Age)
Ind_Age <- 1*(Age >= 6)
S.a <- sens.pcr/(1-sens.pcr)
Sp.a <- spec.pcr/(1-spec.pcr)
  
data <- list(nind=nind,
               y.pcr=y.pcr,
               y.dpp_c=y.dpp_c,
               Sex=Sex2,
               Age=Age,
               h_mu0=h_mu0,
               h_mu1=h_mu1,
               h_p0=h_p0,
               h_p1=h_p1,
               S.a=S.a, ## function of mean sensitivity for PCR
               Sp.a=Sp.a ## function of mean specificity for PCR
             )

## Initialize prior values
inits <- function(){
    list(B0=rnorm(0,1),
         B1=rnorm(0,1),
         B2=rnorm(0,1),
         B3=rnorm(0,1),
         eps=rnorm(nind,0,5000)
         )
}

## MCMC Simulation
sens_spec.sim_contDPP_log <- bugs(data=data, 
                              inits=inits, 
                              model.file="Model_cDPP_mixGaussian.txt", 
                              parameters=c("B0",
                                           "B1",
                                           "B2",
                                           "B3",
                                           "S.pcr",
                                           "Sp.pcr",
                                           "mu1",
                                           "p1",
                                           "mu0",
                                           "p0",
                                           "pi.D",
                                           "D" 
                                           # "eps"
                                           ),
                              n.chains=3, n.iter=30000, n.burnin=10000, 
                              codaPkg=TRUE, debug=TRUE)

## Use coda to read things back into R
codaobject_contDPP_log <- read.bugs(sens_spec.sim_contDPP_log)
model1_contDPP_log <- do.call(rbind.data.frame, codaobject_contDPP_log)

save(model1_contDPP_log, file="G:/My Drive/Research/Sensitivity_Specificity/Models/Model Fits/model2_posterior_log.Rdata")

```

### Results summary (log transformation)



### Table summary

```{r}
tmp2 <- model1_contDPP[,names(model1_contDPP) %in% c("B0","B1","B2","B3","S.pcr","Sp.pcr","a1","b1","a0","b0")]

Pg0_fcn <- function(x){
  mean(x > 0)
}

tmp2_summary <- data.frame(parameter=names(tmp2),
                           mean=apply(tmp2, 2, mean),
                           median=apply(tmp2, 2, median),
                           LB=apply(tmp2, 2, quantile, 0.025),
                           UB=apply(tmp2, 2, quantile, 0.975),
                           Pg0=apply(tmp2, 2, Pg0_fcn))
xtable(tmp2_summary)
```


### Side-by-side boxplot of posterior distributions of parameter values

```{r}
## Collect only population parameters for display in boxplot
tmp <- model1_contDPP[,names(model1_contDPP) %in% c("B0", "B1", "B2", "B3",
                                                    "S.pcr", "Sp.pcr",
                                                    "a1", "b1", "a0", "b0")]
tmp$ID <- 1:nrow(tmp)

tmp_long <- data.frame(variable=rep(c("B0", "B1", "B2", "B3",
                                                    "S.pcr", "Sp.pcr",
                                                    "a1", "b1", "a0", "b0"), 
                                    each=nrow(tmp)),
                       value=c(model1_contDPP$B0,
                                   model1_contDPP$B1,
                                   model1_contDPP$B2,
                                   model1_contDPP$B3,
                                   model1_contDPP$S.pcr,
                                   model1_contDPP$Sp.pcr,
                                   model1_contDPP$a1,
                                   model1_contDPP$b1,
                                   model1_contDPP$a0,
                                   model1_contDPP$b0))

bp1 <- ggplot(data=tmp_long[tmp_long$variable %in% c("B0","B1","B2","B3"),],
              aes(y=value, x=variable)) + 
  geom_boxplot() +
  ylab("Posterior effect estimate") +
  xlab("") + 
  theme_bw()
bp1

bp2 <- ggplot(data=tmp_long[tmp_long$variable %in% c("a1", "b1", "a0", "b0"),],
              aes(y=value, x=variable)) + 
  geom_boxplot() + 
  ylab("Posterior hyperparameter estimates") + 
  xlab("Gamma distribution parameters") + 
  theme_bw()
bp2

bp_3 <- ggplot(data=tmp_long[tmp_long$variable %in% c("S.pcr", "Sp.pcr"),],
               aes(y=value, x=variable)) + 
  geom_boxplot() + 
  ylab("Posterior estimates") + 
  xlab("PCR Sensitivity and Specificity") + 
  theme_bw()
bp_3
```

## Results (with Sqrt transformation)

## References
